---
title: "Cost of living"
author: "Willem"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
> warning=FALSE,error=FALSE,message=FALSE #Could put into setup to make it apply to everything


```{r, echo=TRUE}
library(tidyverse)
library(janitor)
costofliving <- read_csv("costofliving.csv")

costofliving_clean<- costofliving %>%
  clean_names
```

Tidyverse function
- select
- filter
- mutate
- rename
- group_by
- summarize

##Demonstrating select function 
*(which manipulates columns)*
```{r, eval=FALSE} 
costofliving_subset<-costofliving_clean %>%  #This is a comment
  select(cost_of_living_index,groceries_index)
```

##
```{r}
allbutcostofliving<-costofliving_clean %>%
  select(-c(cost_of_living_index, rent_index))
```

```{r}
costofliving_clean %>%
  select(c(-cost_of_living_index))
```


###filter (selects rows)
```{r}
costofliving_clean %>%
  filter(cost_of_living_index<100)
```


```{r}
costofliving_clean %>%
  filter(cost_of_living_index>=100) ## cities that are more or as expensive as New York
```

```{r}
costofliving_clean %>%
  filter(city== "Oslo, Norway") ##Will only give rows where the city is Oslo, Norway
```

```{r}
costofliving_clean %>%
  filter(city %in% c("Oslo, Norway",
                        "Bergen, Norway",
                          "Trondheim, Norway"))
```

### rename

`rename` changes the name of a variable

```{r, eval=FALSE}
dataset %>%
  rename(new_variablr_name = old_variable_name)
```


```{r}
costofliving_clean <- costofliving_clean %>%
  rename(cofi = cost_of_living_index)
```

### separate and unite

```{r}
costofliving_clean <- costofliving_clean %>%
  separate(city, into = c("city", "country"), sep=", ")
```


### mutate
```{r}
costofliving_clean %>%
  select(-c(x1,rank))
```


```{r}
costofliving_clean %>%
  mutate(share = restaurant_price_index/local_purchasing_power_index)
```


##ifelse

```{r}
costofliving_clean %>%
  mutate(norway = ifelse(country %in% c("Norway", "Sweden"), 1, 0))

```

> when multiple values use %in%, when individual ==


### summarise and group_by

1. summarise gives ssummary statistics

- group_by gives the summary statistics per some variable(s)

```{r}
costofliving_clean %>%
  summarise(cofi_sum = sum(cofi, na.rm = TRUE))
```

```{r}
costofliving_clean %>%
  group_by(country) %>%
  summarise(cofi_sum =sum(cofi, na.rm=TRUE)) %>%
ungroup()
```

- min
- max
- mean
- median
- sd
- sum
count

```{r}

costofliving_clean %>%
  group_by(country) %>%
  summarise(avg_cofi = mean(cofi, na.rm=TRUE)) %>%
  ungroup()
```


```{r}
costofliving_clean %>%
  group_by(country) %>%
  count() %>%
  arrange(desc(n))%>%
ungroup()
```
##Nice tables
```{r}
library(kableExtra)

smalltable<-costofliving_clean %>%
  select(country, city, cofi) %>%
  filter(city %in% c("Oslo", "Bergen", "Trondheim"))

smalltable %>%
  kable() %>%
  kable_styling()

```

